---
title: "Moorea LTER"
author: "Carmen Hoyt"
date: last-modified
format:
  html:
      toc: true
      code-fold: true
editor: visual
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

# Moorea Coral Reef LTER 

[Data](https://mcr.lternet.edu/data)

## Data

[benthic water temperature](https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-mcr.1035.16):
- temp (C)
- sites 1-6

[benthic algae cover](https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-mcr.8.36)
- percent cover
- sites 1-6

[Population Dynamics of Acanthaster planci (crown of thorns)](https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-mcr.1039.11)
- counts
- sites 1-6

[corals (stony corals)](https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-mcr.4.40)
- percent cover
- sites 1-6

## Questions:

- How does benthic water temperature relate to benthic algal cover?
- **How does benthic water temperature relate to crown of thorn abundance? By habitat type?**
- How does benthic water temperature relate to stony coral cover?
- How does benthic algal cover relate to crown of thorn abundance? (they eat coral)
- How does benthic algal cover relate to stony coral cover?
- How does stony coral cover relate to crown of thorn abundance?

## Load Packages

```{r}
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

## Load Data

```{r}
#| code-summary: Load Data

# -------- Water Temp --------
# Read in filenames for all LTER sites
water_temp_filenames <- list.files("/Users/choyt/MEDS/eds213-databases/eds-213-discussion/moorrea-LTER-data/benthic-water-temp", full.names=TRUE)

# Create an empty list
all_LTER_sites_water <- vector("list", length = length(water_temp_filenames))

# Loop through filenames
for (i in seq_along(water_temp_filenames)) {
  # Read in files
  df <- read_csv(water_temp_filenames[i], show_col_types = FALSE)
  # Name the df
  df_name <- paste0("LTER0", i-1)
  # Append df to empty list
  all_LTER_sites_water[[i]] <- assign(df_name, df)
}

# Create df for all LTER sites
benthic_water_temp <- bind_rows(all_LTER_sites_water) %>%
  clean_names()

# -------- Crown of Thorns Abundance --------
ac_abundance <- read_csv("/Users/choyt/MEDS/eds213-databases/eds-213-discussion/moorrea-LTER-data/MCR_LTER_COTS_abundance_2005-2022_20220503.csv", show_col_types = FALSE) %>%
  clean_names()
```

## Clean Data

```{r}
# Clean water temp data
benthic_water_temp_cleaned <- benthic_water_temp %>%
  # Extract year from time_local
  mutate(time_local = as_date(time_local)) %>%
  mutate(year = year(time_local)) %>%
  # Find average water temp by site and year
  group_by(year, site) %>%
  summarise(avg_year_temp_c = mean(temperature_c))

# Clean ac abundance data
ac_abundance_cleaned <- ac_abundance %>%
  # Convert site to match water temp
  mutate(site = case_when(
    site == "LTER 1" ~ "LTER01",
    site == "LTER 2" ~ "LTER02",
    site == "LTER 3" ~ "LTER03",
    site == "LTER 4" ~ "LTER04",
    site == "LTER 5" ~ "LTER05",
    site == "LTER 6" ~ "LTER06"
  )) %>%
  rename(a_planci_counts = cots)%>%
  group_by(year, site) %>%
  summarise(a_planci_counts = sum(a_planci_counts, na.rm = TRUE))

write_csv(benthic_water_temp_cleaned, "/Users/choyt/MEDS/eds213-databases/eds-213-discussion/moorrea-LTER-data/benthic_water_temp_cleaned.csv")

write_csv(ac_abundance_cleaned, "/Users/choyt/MEDS/eds213-databases/eds-213-discussion/moorrea-LTER-data/ac_abundance_cleaned.csv")
```

## Join Data

```{r}
# Left join water temp to ac abundance 
#temp_counts_joined <- left_join(ac_abundance_cleaned, benthic_water_temp_cleaned, by = c("site", "year"))
```

## Create Database

.open moorea-LTER.db
CREATE TABLE Abundance (
  year INT PRIMARY KEY
  site VARCHAR PRIMARY KEY
  UNIQUE (year, site)
  habitat VARCHAR
  transect INT
  a_planci_counts INT
);

COPY Abundance FROM ac_abundance_cleaned (header TRUE);

CREATE TABLE Water_temp (
  year INT
  site VARCHAR
  UNIQUE (year, site)
  avg_year_temp_c REAL
);

COPY Water_temp FROM benthic_water_temp_cleaned (header TRUE);
